---
title: "Solutions HW#4 SOC 574 - Joseph Risi"
output: word_document
---

```{r, echo = F, warning = F, message = F}
library(here)
library(purrr)
library(dplyr)
library(tidyr)
library(forcats)
library(ggplot2)
library(flextable)
```

```{r, echo = F}
data <- foreign::read.dta(here("data", "nlsy.dta"))
```

```{r, echo = F, message = F, warning = F}
data_split <-
    data %>%
    group_by(int1wksc) %>%
    summarise(mean = mean(grades8),
              sd = sd(grades8),
              n = n(),
              stdError = sd / sqrt(n),
              tscore_95 = qt(1 - 0.05 / 2, df = n - 1),
              tscore_98 = qt(1 - 0.02 / 2, df = n - 1),
              marginOfError_95 = tscore_95 * stdError,
              marginOfError_98 = tscore_98 * stdError,
              lowerCI_95 = mean - marginOfError_95,
              upperCI_95 = mean + marginOfError_95,
              lowerCI_98 = mean - marginOfError_98,
              upperCI_98 = mean + marginOfError_98) %>%
    pivot_longer(cols = !matches("^mean$|^sd$|^n$|^stdError$|int1wksc"),
                 names_to = c("variables", "confidence"),
                 names_pattern = "([a-zA-Z]{1,13})_([0-9]{2})") %>%
    pivot_wider(names_from = "variables", values_from = "value") %>%
    mutate(ci = paste0("(", round(lowerCI, 4), ", ", round(upperCI, 4), ")"),
              .keep = "unused") %>%
    arrange(confidence)
```

```{r, echo = F}
flextable(data_split)
```

**3a)**

The 95% confidence interval for those students who engaged in intensive work during high school is approximately `r data_split %>% filter(int1wksc == 1, confidence == 95) %>% pull(ci)`. Of the students who engaged in intensive work, I am 95% certain their true **average** grade score lies between `r data_split %>% filter(int1wksc == 1, confidence == 95) %>% pull(ci)` with my best estimate being `r data_split %>% filter(int1wksc == 1, confidence == 95) %>% pull(mean)`.

**3b)**

The 95% confidence interval for those students who **did not engage** in intensive work during high school is approximately `r data_split %>% filter(int1wksc == 0, confidence == 95) %>% pull(ci)`. Of the students who did not engage in intensive work, I am 95% certain their true **average** grade score lies between `r data_split %>% filter(int1wksc == 0, confidence == 95) %>% pull(ci)` with my best estimate being `r data_split %>% filter(int1wksc == 0, confidence == 95) %>% pull(mean)`.

**3c)**

The 95% confidence intervals do slightly overlap for the two groups which means I am not convinced that intensive work negatively affects grades.

**3d)**

The intervals got wider since I made my t-scores higher. The qualitative judgment concerning the effect of intensive work on grades remains the same, however. Meaning, I am still not convinced intensive work has a negative impact on grades.