---
title: 'Sarah & Joe Present: Does time spent doing homework in adolescence increasing reading and math achievement?'
output: pdf_document
---

## Packages

```{r, warning = F, message = F}
library(mlr)
library(here)
library(purrr)
library(dplyr)
library(tidyr)
library(haven)
library(forcats)
library(ggplot2)
library(stringr)
library(sjlabelled)
```

## Read in data and filter out all those who were not in the wave

```{r}
data <- read_dta(here("class data.dta"))

baseYearCohort <-
    data %>%
    select(matches("by|sex|race|^id$")) %>%
    filter(byhomewk != 99)
```

Number of students removed due to not being in the wave: `r nrow(data) - nrow(baseYearCohort)`  
Fraction of total students removed: `r (nrow(data) - nrow(baseYearCohort)) / nrow(data)`

## Get a sense for the data

```{r}
ggplot(baseYearCohort, aes(x = as_label(byhomewk))) +
    geom_bar() +
    theme_bw() +
    coord_flip() +
    labs(x = "Amount of time spent doing homework (hours)")

ggplot(baseYearCohort, aes(x = as.numeric(by2xrirr))) +
    geom_histogram(bins = 35, color = "red") +
    labs(x = "Reading Test Scores") +
    theme_bw()

ggplot(baseYearCohort, aes(x = as.numeric(by2xmirr))) +
    geom_histogram(bins = 35, color = "red") +
    labs(x = "Math Test Scores") +
    theme_bw()

ggplot(baseYearCohort, aes(x = as_label(byhomewk), y = as.numeric(by2xrirr))) +
    geom_boxplot() +
    theme_bw() +
    coord_flip() +
    labs(x = "Amount of time doing homework (hours)", y = "Reading Test Scores")

ggplot(baseYearCohort, aes(x = as_label(byhomewk), y = as.numeric(by2xmirr))) +
    geom_boxplot() +
    theme_bw() +
    coord_flip() +
    labs(x = "Amount of time doing homework (hours)", y = "Math Test Scores")

ggplot(baseYearCohort, aes(x = as.numeric(by2xrirr), y = as.numeric(by2xmirr))) +
    labs(x = "Reading Test Score", y = "Math Test Score") +
    geom_point() +
    theme_bw()

cor(baseYearCohort$by2xrirr, baseYearCohort$by2xmirr, use = "complete.obs")
```

## Let's run some basic regressions

```{r}
baseYearCohortWide <-
    baseYearCohort %>%
    mutate(byhomewk_label = as_label(byhomewk),
           byhomewk = as.numeric(as.logical(as.numeric(byhomewk)))) %>%
    pivot_wider(names_from = byhomewk_label,
                values_from = byhomewk,
                values_fill = F)

readingTestRegression <-
    lm(by2xrirr ~ `.50 to 1.99 hours` + `2.00 to 2.99 hours` +
           `3.00 to 5.49 hours` + `5.50 to 10.49 hours` +
           `10.50 to 12.99 hours` + `13.00 to 20.99 hours` +
           `21.00 and up hours`,
       data = baseYearCohortWide)
summary(readingTestRegression)

mathTestRegression <-
    lm(by2xmirr ~ `.50 to 1.99 hours` + `2.00 to 2.99 hours` +
           `3.00 to 5.49 hours` + `5.50 to 10.49 hours` +
           `10.50 to 12.99 hours` + `13.00 to 20.99 hours` +
           `21.00 and up hours`,
       data = baseYearCohortWide)
summary(mathTestRegression)
```

## Let's throw in some controls and make these regressions a bit more complicated

```{r}
CreateWideControls <- function(df, test) {
    df %>%
        mutate(across(c(byhomewk, sex, race, bypared, byfaminc),
                      list(label = compose(as.character, as_label))),
               by2xrirr = as.numeric(by2xrirr),
               by2xmirr = as.numeric(by2xmirr)) %>%
        select(matches(paste0("label|^id$|", test))) %>%
        filter(across(everything(), ~!is.na(.x))) %>%
        mutate(byfaminc_label = if_else(byfaminc_label == "none",
                                        "$0",
                                        byfaminc_label)) %>%
        mutate(across(c(byhomewk_label, sex_label, race_label, bypared_label,
                        byfaminc_label),
                      as.factor))
}

RelocatePretty <- function(df) {
    df %>%
        relocate(".50.to.1.99.hours", "2.00.to.2.99.hours",
                 "3.00.to.5.49.hours", "5.50.to.10.49.hours",
                 "10.50.to.12.99.hours", "13.00.to.20.99.hours",
                 "21.00.and.up.hours", "less.than..1.000",
                 ".1.000....2.999", ".3.000....4.999", ".5.000....7.499",
                 ".7.500....9.999", ".10.000..14.999", ".15.000..19.999",
                 ".20.000..24.999", ".25.000..34.999", ".35.000..49.999",
                 ".50.000..74.999", ".75.000..99.999", ".100.000.199.999",
                 ".200.000.or.more", "..hs.....4yr.deg", "h.s..grad.or.ged",
                 "college.graduate", "m.a..equivalent", "ph.d...m.d...other") %>%
        select(-id, -black.not.hispanic, -male, -none, -didn.t.finish.hs, -`.0`)
}

# Reading Scores
readScoresWide <-
    CreateWideControls(baseYearCohort, "by2xrirr") %>%
    createDummyFeatures()
    
colnames(readScoresWide) <- str_replace(colnames(readScoresWide),
                                        "by.*_label.|race_label.|sex_label.",
                                        "")

readScoresWide <- RelocatePretty(readScoresWide)

# Math Scores
mathScoresWide <-
    CreateWideControls(baseYearCohort, "by2xmirr") %>%
    createDummyFeatures()

colnames(mathScoresWide) <- str_replace(colnames(mathScoresWide),
                                        "by.*_label.|race_label.|sex_label.",
                                        "")

mathScoresWide <- RelocatePretty(mathScoresWide)
```

Number of rows removed due to missing values in the control variables: `r nrow(baseYearCohort) - nrow(readScoresWide)`  
Fraction removed: `r (nrow(baseYearCohort) - nrow(readScoresWide)) / nrow(baseYearCohort)`

The math and reading scores differ by 1 in terms of missing values.

```{r}
readingTestRegressionControls <- lm(by2xrirr ~ ., data = readScoresWide)
summary(readingTestRegressionControls)

mathTestRegressionControls <- lm(by2xmirr ~ ., data = mathScoresWide)
summary(mathTestRegressionControls)
```

## Let's recode by family income

```{r}
ggplot(baseYearCohort, aes(x = as_label(byfaminc))) +
    geom_bar() +
    theme_bw() +
    coord_flip() +
    labs(x = "Family Income")

baseYearCohortRecode <-
    baseYearCohort %>%
    mutate(byfaminc = as.character(as_label(byfaminc)),
           byfaminc = fct_recode(byfaminc,
                                 "Less than $10,000" = "none",
                                 "Less than $10,000" = "less than $1,000",
                                 "Less than $10,000" = "$1,000 - $2,999",
                                 "Less than $10,000" = "$3,000 - $4,999",
                                 "Less than $10,000"= "$5,000 - $7,499",
                                 "Less than $10,000" = "$7,500 - $9,999",
                                 "$10,000 - $19,999" = "$10,000-$14,999",
                                 "$10,000 - $19,999" = "$15,000-$19,999",
                                 "$75,000 and above" = "$75,000-$99,999",
                                 "$75,000 and above" = "$100,000-199,999",
                                 "$75,000 and above" = "$200,000 or more"),
           byfaminc = fct_relevel(byfaminc,
                                  "Less than $10,000",
                                  "$10,000 - $19,999",
                                  "$20,000-$24,999",
                                  "$25,000-$34,999",
                                  "$35,000-$49,999",
                                  "$50,000-$74,999",
                                  "$75,000 and above"))

ggplot(baseYearCohortRecode, aes(x = byfaminc)) +
    geom_bar() +
    theme_bw() +
    coord_flip() +
    labs(x = "Family Income")

CreateWideControls <- function(df, test) {
    df %>%
        mutate(across(c(byhomewk, sex, race, bypared, byfaminc),
                      list(label = compose(as.character, as_label))),
               by2xrirr = as.numeric(by2xrirr),
               by2xmirr = as.numeric(by2xmirr)) %>%
        select(matches(paste0("label|^id$|", test))) %>%
        filter(across(everything(), ~!is.na(.x))) %>%
        mutate(byfaminc_label = if_else(byfaminc_label == "none",
                                        "$0",
                                        byfaminc_label)) %>%
        mutate(across(c(byhomewk_label, sex_label, race_label, bypared_label,
                        byfaminc_label),
                      as.factor))
}

RelocatePrettyRecode <- function(df) {
    df %>%
        relocate(".50.to.1.99.hours", "2.00.to.2.99.hours",
                 "3.00.to.5.49.hours", "5.50.to.10.49.hours",
                 "10.50.to.12.99.hours", "13.00.to.20.99.hours",
                 "21.00.and.up.hours", ".10.000....19.999", ".20.000..24.999",
                 ".25.000..34.999", ".35.000..49.999", ".50.000..74.999",
                 ".75.000.and.above", "..hs.....4yr.deg", "h.s..grad.or.ged",
                 "college.graduate", "m.a..equivalent", "ph.d...m.d...other") %>%
        select(-id, -black.not.hispanic, -male, -none, -didn.t.finish.hs,
               -`Less.than..10.000`)
}

# Reading Scores
readScoresWideRecode <-
    CreateWideControls(baseYearCohortRecode, "by2xrirr") %>%
    createDummyFeatures()
    
colnames(readScoresWideRecode) <- str_replace(colnames(readScoresWideRecode),
                                        "by.*_label.|race_label.|sex_label.",
                                        "")

readScoresWideRecode <- RelocatePrettyRecode(readScoresWideRecode)

readingTestRegressionControlsRecode <- lm(by2xrirr ~ .,
                                          data = readScoresWideRecode)
summary(readingTestRegressionControlsRecode)

# Math Scores
mathScoresWideRecode <-
    CreateWideControls(baseYearCohortRecode, "by2xmirr") %>%
    createDummyFeatures()
    
colnames(mathScoresWideRecode) <- str_replace(colnames(mathScoresWideRecode),
                                        "by.*_label.|race_label.|sex_label.",
                                        "")

mathScoresWideRecode <- RelocatePrettyRecode(mathScoresWideRecode)

mathTestRegressionControlsRecode <- lm(by2xmirr ~ ., data = mathScoresWideRecode)
summary(mathTestRegressionControlsRecode)
```
