---
title: "Week 2"
output: pdf_document
---

## Packages

```{r, warning = F, message = F}
library(here)
library(dplyr)
library(purrr)
library(haven)
library(reghelper)
library(sjlabelled)
```

## Data

```{r}
data <- read_dta(here("class data.dta"))
filterData <- filter(data, !is.na(bypared), !is.na(bygrads), !is.na(bys42a))
```

## Correlational Matrix and Descriptive Statistics (naive since we're working wiith categorical variables)

```{r}
cor(dplyr::select(filterData, bygrads, bys42a, bypared), use = "complete.obs")
map(dplyr::select(filterData, bygrads, bys42a, bypared), summary)
```

## Regression

```{r}
model <- lm(bygrads ~ bys42a + bypared, data = filterData)
summary(model)

# Standardizing coefficients

# For every unit increase in x, you get a std. dev. increase in y
beta(model, x = F)

# For every std. dev. increase in x, you get "coefficient" increase in y
beta(model, y = F)

# For every std. dev. increase in x, you get std. dev. increase in y
beta(model)
```

For predicting values, you plug in the values of interest and multiply by the coefficients and add the intercept. To hold a variable constant, you use the mean value of that variable.

So below, we're getting the expected values of your GPA for each value of TV watching while holding parents' education constant.

## Prediction

```{r}
newData <- tibble(bys42a = sort(unique(filterData$bys42a)),
                  bypared = rep(mean(filterData$bypared), length(bys42a)))

predict(model, newdata = newData,
        se.fit = T,
        interval = "confidence",
        level = 0.95)
```
