---
title: "Week 8 Class Notes"
output: html_document
---

https://stats.idre.ucla.edu/r/dae/multinomial-logistic-regression/

* You can interpret using log of the odds.
* A good way to think of multinomial logit models is that you are simultaneously estimating binary logits for all comparisons on your nominal outcome.

```{r, warning = F, message = F}
library(car)
library(here)
library(nnet)
library(dplyr)
library(tidyr)
library(purrr)
library(haven)
library(broom)
library(DAMisc)
library(lmtest)
library(mlogit)
library(mnlogit)
library(effects)
library(forcats)
library(ggplot2)
library(sjlabelled)
```

```{r}
data <- haven::read_dta(here("class data.dta"))
```

```{r}
dataClean <-
    data %>%
    select(sex, bys55f, by2xrirr, byfaminc, matches("f2s30")) %>%
    mutate(spt1 = as.character(sjlabelled::as_factor(f2s30aa)),
           spt2 = as.character(sjlabelled::as_factor(f2s30ab)),
           spt3 =as.character(sjlabelled::as_factor(f2s30bj)),
           spt4 = as.character(sjlabelled::as_factor(f2s30bk)),
           ex1 = as.character(sjlabelled::as_factor(f2s30ac)),
           ex2 = as.character(sjlabelled::as_factor(f2s30ba)),
           ex3 = as.character(sjlabelled::as_factor(f2s30bb)),
           ex4 = as.character(sjlabelled::as_factor(f2s30bc)),
           ex5 = as.character(sjlabelled::as_factor(f2s30bd)),
           ex6 = as.character(sjlabelled::as_factor(f2s30be)),
           ex7 = as.character(sjlabelled::as_factor(f2s30bf)),
           ex8 = as.character(sjlabelled::as_factor(f2s30bg)),
           ex9 = as.character(sjlabelled::as_factor(f2s30bh)),
           ex10 = as.character(sjlabelled::as_factor(f2s30bi))) %>%
    mutate(across(matches("spt[0-9]|ex[0-9]{1,2}"), ~na_if(.x, "1"))) %>%
    mutate(across(matches("spt[0-9]|ex[0-9]{1,2}"), ~replace_na(.x, ""))) %>%
    mutate(across(matches("spt[0-9]|ex[0-9]{1,2}"),
                  ~as.character(fct_recode(.x,
                                           "0" = "2", "1" = "3", "1" = "4", "1" = "5")))) %>%
    rowwise() %>%
    mutate(sport = max(spt1, spt2, spt3, spt4),
           extra = max(ex1, ex2, ex3, ex4, ex5, ex6, ex7, ex8, ex9, ex10),
           extra12 = case_when(sport == "1" & extra == "1" ~ "both",
                               sport == "0" & extra == "0" ~ "noextra",
                               sport == "1" & extra != "" ~ "sportonly",
                               sport != "" & extra == "1" ~ "nonsportonly")) %>%
    select(sex, bys55f, by2xrirr, byfaminc, extra12) %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(sex = as.numeric(if_else(sex == 2, 0, as.numeric(sex))),
           bys55f = as.numeric(bys55f),
           by2xrirr = as.numeric(by2xrirr),
           byfaminc = as.numeric(byfaminc),
           extra12 = factor(extra12))
```

```{r}
models <- map(list("nonsportonly", "noextra", "sportonly", "both"),
              function(df, level) {
                  df$extra12 <- relevel(df$extra12, ref = level)
                  nnet::multinom(extra12 ~ ., data = df)},
              df = dataClean)

map(models, summary)

pp <- fitted(models[[1]])
pp2 <- fitted(models[[2]])

avgp <- Effect(focal.predictors = c("bys55f", "sex", "byfaminc", "by2xrirr"),
               models[[1]])

modelPrediction <- pp %>% as_tibble() %>% bind_cols(dataClean)
```

```{r}
maleVsFemale <- tibble(sex = c(0, 1),
                       byfaminc = mean(dataClean$byfaminc),
                       bys55f = mean(dataClean$bys55f),
                       by2xrirr = mean(dataClean$by2xrirr))

ppSex <-
    predict(models[[1]], newdata = maleVsFemale, type = "probs") %>%
    as_tibble() %>%
    bind_cols(maleVsFemale)

fighters <- tibble(sex = mean(dataClean$sex),
                   byfaminc = mean(dataClean$byfaminc),
                   bys55f = c(0, 1, 2),
                   by2xrirr = mean(dataClean$by2xrirr))

ppFight <-
    predict(models[[1]], newdata = fighters, type = "probs") %>%
    as_tibble() %>%
    bind_cols(fighters)

readScores <- tibble(sex = mean(dataClean$sex),
                     byfaminc = mean(dataClean$byfaminc),
                     bys55f = mean(dataClean$bys55f),
                     by2xrirr = unique(dataClean$by2xrirr))

ppReadScores <-
    predict(models[[1]], newdata = readScores, type = "probs") %>%
    as_tibble() %>%
    bind_cols(readScores) %>%
    pivot_longer(-c(sex, byfaminc, bys55f, by2xrirr))

ggplot(ppReadScores, aes(x = by2xrirr, y = value)) +
    geom_line(aes(group = name, color = name)) +
    theme_bw()

byFamInc <- tibble(sex = mean(dataClean$sex),
                   byfaminc = unique(dataClean$byfaminc),
                   bys55f = mean(dataClean$bys55f),
                   by2xrirr = mean(dataClean$by2xrirr))

ppFamInc <-
    predict(models[[1]], newdata = byFamInc, type = "probs") %>%
    as_tibble() %>%
    bind_cols(byFamInc) %>%
    pivot_longer(-c(sex, byfaminc, bys55f, by2xrirr))

ggplot(ppFamInc, aes(x = byfaminc, y = value)) +
    geom_point(aes(color = name)) +
    geom_line(aes(group = name, color = name)) +
    theme_bw()
```

Do your variables affect ANY of the outcomes? To test this, you hold out the variable and re-estimate the model and see if the variance explained changes in a significant fashion.

```{r}
# Hard Way
model <- nnet::multinom(extra12 ~ ., data = dataClean)
lrtest(model, "bys55f")
lrtest(model, "byfaminc")
lrtest(model, "sex")
lrtest(model, "by2xrirr")

# Easier Way, don't think I can do a Wald Test.
# Well I think I could, would need to use mlogit or mnlogit
modelDf <- broom::tidy(models[[1]], exponentiate = T)
car::Anova(models[[1]])
```

Do your variables affect these specific two outcomes? Can my predictor distinguish between categories? If not, then they should be combined.

```{r}
# Appears to be using Wald Test for combining alternatives
# I think I'd have to implement combination test manually for LR
combTest(models[[1]])
```

https://stats.idre.ucla.edu/r/dae/multinomial-logistic-regression/
https://bookdown.org/chua/ber642_advanced_regression/multinomial-logistic-regression.html