---
title: "Presentation with Anna Shetler & Joe Risi"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = F)
```

Seminar leadership question: Does watching too much television make you feel useless?

**Presentation Need to Haves**
* Consistent sample (Same N for everything, deal with missing)
* Descriptive Graphs and Tables
    * Bar Chart, Histogram, Violin Plots, Mean/Median/Std. Dev. for each variable.
* Multicollinearity Check
* Regression with just amount of TV watched (weekend + weekday) for by and f2.
    * Qualitatively different to be watching a lot of TV during the week vs. the weekend.
* 2nd regression with everything else (controls + confounders) for by and f2.
* ANOVA test to compare models 1 + 2.

**Strategy**

1. bys42a (amount of TV watched during the weekday)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 4 hours
    * 4 - 5 hours
    * over 5 hours a day
    
2. bys42b (amount of TV watched during the weekend)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 4 hours
    * 4 - 5 hours
    * over 5 hours a day

3. f2s35a (amount of TV watched during weekdays)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 5 hours
    * over 5 hours a day
    
4. f2s35b (amount of TV watched during weekends)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 5 hours
    * over 5 hours a day

5. bys44i (I certainly feel useless at times)
    * strongly agree
    * agree
    * disagree
    * strongly disagree
    
6. f2s66i (r feels useless at times)
    * strongly agree
    * agree
    * disagree
    * strongly disagree

Confounders (bys):
    * teachers praise my efforts (bys59h)
        * strongly agree
        * agree
        * disagree
        * strongly disagree
        
Confounders (f2)
    * f2s81c (last 30 days, # times r drank alcohol)
        * 0 occasions
        * 1-2 occcasions
        * 3-19 occasions
        * 20+ occasions
    * f2s82 (# times r had 5 drinks or more in a row)
        * none
        * once
        * twice
        * 3-5 times
        * 6-9 times
        * 10 or more times

Controls:
    * sex (sex)
    * race (race)
    * parent's highest level of education (bypared)
    * family income (byfaminc)

## Packages

```{r, message = F, warning = F}
library(mlr)
library(here)
library(MASS)
library(dplyr)
library(haven)
library(forcats)
library(ggplot2)
library(stringr)
library(sjlabelled)
library(ggcorrplot)
```

## Read in data and clean data

```{r}
data <- haven::read_dta(here("class data.dta"))

baseYear <- select(data, bys42a, bys42b, bys44i, sex, race, bypared, byfaminc)
followYear <- select(data, f2s35a, f2s35b, f2s66i, sex, race, bypared, byfaminc)

baseYearClean <-
    baseYear %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(bys44i = as.character(fct_recode(sjlabelled::as_factor(bys44i),
                                            "1" = "4",
                                            "2" = "3",
                                            "3" = "2",
                                            "4" = "1")))

followYearClean <-
    followYear %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(f2s66i = as.character(fct_recode(sjlabelled::as_factor(f2s66i),
                                            "1" = "4",
                                            "2" = "3",
                                            "3" = "2",
                                            "4" = "1")))
```

## Correlation Matrices, Nothing looks too correlated thankfully

```{r}
byCorr <- cor(select(baseYearClean, -sex, -race))
fyCorr <- cor(select(followYearClean, -sex, -race))

ggcorrplot(byCorr,
           type = "lower",
           hc.order = T,
           method = "circle",
           colors = c("red", "white", "blue")) +
    ggtitle("Correlational Matrix For Base Year")

ggcorrplot(fyCorr,
           type = "lower",
           hc.order = T,
           method = "circle",
           colors = c("red", "white", "blue")) +
    ggtitle("Correlational Matrix For Follow-Up Year")
```

## Run Ordinal Logistic Regression

```{r}
baseYearModel1Wide <-
    baseYearClean %>%
    select(bys44i, bys42a, bys42b) %>%
    mutate(bys42a = as.factor(as_character(bys42a)),
           bys42b = as.factor(as_character(bys42b))) %>%
    createDummyFeatures() %>%
    relocate("bys42a.don.t.watch.tv", "bys42a.lt.1.hour.a.day",
             "bys42a.1.2.hours", "bys42a.2.3.hours", "bys42a.3.4.hours",
             "bys42a.4.5.hours", "bys42a.over.5.hrs.a.day",
             "bys42b.don.t.watch.tv", "bys42b.lt.1.hour.a.day",
             "bys42b.1.2.hours", "bys42b.2.3.hours", "bys42b.3.4.hours",
             "bys42b.4.5.hours", "bys42b.over.5.hrs.a.day") %>%
    mutate(bys44i = as.factor(bys44i)) %>%
    select(-bys42a.don.t.watch.tv, -bys42b.don.t.watch.tv)

ordinalBasicBase <- MASS::polr(bys44i ~ ., data = baseYearModel1Wide)
```
