---
title: "Presentation with Anna Shetler & Joe Risi"
output: pdf_document
---

```{r, echo = F}
knitr::opts_chunk$set(echo = F)
```

Seminar leadership question: Does watching too much television make you feel useless?  

```{r, message = F, warning = F}
library(mlr)
library(car)
library(here)
library(MASS)
library(dplyr)
library(haven)
library(brant)
library(broom)
library(forcats)
library(ggplot2)
library(stringr)
library(sjlabelled)
library(ggcorrplot)
```

```{r}
data <- haven::read_dta(here("class data.dta"))
dataClean <-
    data %>%
    mutate(byfaminc = as_character(byfaminc),
           byfaminc = fct_recode(byfaminc,
                                 "Less than $10,000" = "none",
                                 "Less than $10,000" = "less than $1,000",
                                 "Less than $10,000" = "$1,000 - $2,999",
                                 "Less than $10,000" = "$3,000 - $4,999",
                                 "Less than $10,000"= "$5,000 - $7,499",
                                 "Less than $10,000" = "$7,500 - $9,999",
                                 "$10,000 - $19,999" = "$10,000-$14,999",
                                 "$10,000 - $19,999" = "$15,000-$19,999",
                                 "$75,000 and above" = "$75,000-$99,999",
                                 "$75,000 and above" = "$100,000-199,999",
                                 "$75,000 and above" = "$200,000 or more"),
           byfaminc = fct_relevel(byfaminc,
                                  "Less than $10,000",
                                  "$10,000 - $19,999",
                                  "$20,000-$24,999",
                                  "$25,000-$34,999",
                                  "$35,000-$49,999",
                                  "$50,000-$74,999",
                                  "$75,000 and above"),
           sex = as.factor(as_character(sex)),
           race = as.factor(as_character(race)),
           bypared = as.factor(as_character(bypared)))

baseYear <- select(dataClean, bys42a, bys42b, bys44i, sex, race, bypared, byfaminc)
followYear <- select(dataClean, f2s35a, f2s35b, f2s66i, sex, race, bypared, byfaminc)

baseYearClean <-
    baseYear %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(bys44i = as.character(fct_recode(sjlabelled::as_factor(bys44i),
                                            "1" = "4",
                                            "2" = "3",
                                            "3" = "2",
                                            "4" = "1")))

followYearClean <-
    followYear %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(f2s66i = as.character(fct_recode(sjlabelled::as_factor(f2s66i),
                                            "1" = "4",
                                            "2" = "3",
                                            "3" = "2",
                                            "4" = "1")))
```

## Correlation Matrices, Nothing looks too correlated thankfully

```{r}
baseYearCorr <-
    baseYearClean %>%
    mutate(bys44i = as.numeric(bys44i),
           bys42a = as.numeric(bys42a),
           bys42b = as.numeric(bys42b),
           bypared = as.numeric(bypared),
           byfaminc = as.numeric(byfaminc))

followYearCorr <-
    followYearClean %>%
    mutate(f2s66i = as.numeric(f2s66i),
           f2s35a = as.numeric(f2s35a),
           f2s35b = as.numeric(f2s35b),
           bypared = as.numeric(bypared),
           byfaminc = as.numeric(byfaminc))

byCorr <- cor(select(baseYearCorr, -sex, -race))
fyCorr <- cor(select(followYearCorr, -sex, -race))

ggcorrplot(byCorr,
           type = "lower",
           method = "circle",
           colors = c("red", "white", "blue")) +
    ggtitle("Correlational Matrix For Base Year")

ggcorrplot(fyCorr,
           type = "lower",
           method = "circle",
           colors = c("red", "white", "blue")) +
    ggtitle("Correlational Matrix For Follow-Up Year")
```

## Run Ordinal Logistic Regression

### Run the model for the base year only including TV watching

```{r}
baseYearModel <-
    baseYearClean %>%
    select(bys44i, bys42a, bys42b) %>%
    mutate(bys44i = as.factor(bys44i),
           bys42a = as.numeric(bys42a),
           bys42b = as.numeric(bys42b))

ordinalBasicBase <- MASS::polr(bys44i ~ ., data = baseYearModel)
summary(ordinalBasicBase)
exp(coef(ordinalBasicBase))
obb <- broom::tidy(ordinalBasicBase, conf.int = T, exponentiate = T, p.value = T)
car::poTest(ordinalBasicBase)
```

### Run the model for the base year only including TV watching as a categorical varible

```{r}
baseYearModelWide <-
    baseYearClean %>%
    select(bys44i, bys42a, bys42b) %>%
    mutate(bys42a = as.factor(as_character(bys42a)),
           bys42b = as.factor(as_character(bys42b))) %>%
    createDummyFeatures() %>%
    relocate("bys42a.don.t.watch.tv", "bys42a.lt.1.hour.a.day",
             "bys42a.1.2.hours", "bys42a.2.3.hours", "bys42a.3.4.hours",
             "bys42a.4.5.hours", "bys42a.over.5.hrs.a.day",
             "bys42b.don.t.watch.tv", "bys42b.lt.1.hour.a.day",
             "bys42b.1.2.hours", "bys42b.2.3.hours", "bys42b.3.4.hours",
             "bys42b.4.5.hours", "bys42b.over.5.hrs.a.day") %>%
    mutate(bys44i = as.factor(bys44i)) %>%
    select(-bys42a.over.5.hrs.a.day, -bys42b.over.5.hrs.a.day)

ordinalBasicBaseWide <- MASS::polr(bys44i ~ ., data = baseYearModelWide)
summary(ordinalBasicBaseWide)
exp(coef(ordinalBasicBaseWide))
```

### Run the model for the base year only with TV watching and controls

```{r}
baseYearModelControls <-
    baseYearClean %>%
    createDummyFeatures() %>%
    relocate("bypared.didn.t.finish.hs", "bypared.h.s..grad.or.ged",
             "bypared...hs.....4yr.deg", "bypared.college.graduate",
             "bypared.m.a..equivalent", "bypared.ph.d...m.d...other") %>%
    mutate(bys44i = as.factor(bys44i),
           bys42a = as.numeric(bys42a),
           bys42b = as.numeric(bys42b)) %>%
    select(-sex.male, -race.white.not.hispanic, -byfaminc.Less.than..10.000,
           -bypared.didn.t.finish.hs)

ordinalControlsBase <- MASS::polr(bys44i ~ ., data = baseYearModelControls)
summary(ordinalControlsBase)
exp(coef(ordinalControlsBase))
brant(ordinalControlsBase)
```

### Run the model for the follow-up year only including TV watching

```{r}
followYearModel <-
    followYearClean %>%
    select(f2s66i, f2s35a, f2s35b) %>%
    mutate(f2s66i = as.factor(f2s66i),
           f2s35a = as.numeric(f2s35a),
           f2s35b = as.numeric(f2s35b))

ordinalBasicFollow <- MASS::polr(f2s66i ~ ., data = followYearModel)
summary(ordinalBasicFollow)
exp(coef(ordinalBasicFollow))
brant(ordinalBasicFollow)
```

### Run the model for the follow-up year only including TV watching as a categorical variable

```{r}
followYearModelWide <-
    followYearClean %>%
    select(f2s66i, f2s35a, f2s35b) %>%
    mutate(f2s35a = as.factor(as_character(f2s35a)),
           f2s35b = as.factor(as_character(f2s35b))) %>%
    createDummyFeatures() %>%
    relocate("f2s35a.don.t.watch.tv", "f2s35a.less.1hr.day",
             "f2s35a.1.2.hours.day", "f2s35a.2.3.hours.day", "f2s35a.3.5.hours.day",
             "f2s35a.5.hrs.more.a.day",
             "f2s35b.don.t.watch.tv", "f2s35b.less.1hr.day",
             "f2s35b.1.2.hours.day", "f2s35b.2.3.hours.day", "f2s35b.3.5.hours.day",
             "f2s35b.5hrs.more.a.day") %>%
    mutate(f2s66i = as.factor(f2s66i)) %>%
    select(-f2s35a.5.hrs.more.a.day, -f2s35b.5hrs.more.a.day)

ordinalBasicFollowWide <- MASS::polr(f2s66i ~ ., data = followYearModelWide)
summary(ordinalBasicFollowWide)
exp(coef(ordinalBasicFollowWide))
brant(ordinalBasicFollowWide)
```

### Run the model for the follow-up year only including TV watching + controls

```{r}
followYearModelControls <-
    followYearClean %>%
    createDummyFeatures() %>%
    relocate("bypared.didn.t.finish.hs", "bypared.h.s..grad.or.ged",
             "bypared...hs.....4yr.deg", "bypared.college.graduate",
             "bypared.m.a..equivalent", "bypared.ph.d...m.d...other") %>%
    mutate(f2s66i = as.factor(f2s66i),
           f2s35a = as.numeric(f2s35a),
           f2s35b = as.numeric(f2s35b)) %>%
    select(-sex.male, -race.white.not.hispanic, -byfaminc.Less.than..10.000,
           -bypared.didn.t.finish.hs)

ordinalControlsFollow <- MASS::polr(f2s66i ~ ., data = followYearModelControls)
summary(ordinalControlsFollow)
exp(coef(ordinalControlsFollow))
brant(ordinalControlsFollow)
```
