---
title: "Presentation with Anna Shetler & Joe Risi"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = F)
```

Seminar leadership question: Does watching too much television make you feel useless?

**Presentation Need to Haves**
* Consistent sample (Same N for everything, deal with missing)
* Descriptive Graphs and Tables
    * Bar Chart, Histogram, Violin Plots, Mean/Median/Std. Dev. for each variable.
* Regression with just amount of TV watched (weekend + weekday) for by and f2.
    * Qualitatively different to be watching a lot of TV during the week vs. the weekend.
* 2nd regression with everything else (controls + confounders) for by and f2.
* ANOVA test to compare models 1 + 2.

**Strategy**

1. bys42a (amount of TV watched during the weekday)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 4 hours
    * 4 - 5 hours
    * over 5 hours a day
    
2. bys42b (amount of TV watched during the weekend)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 4 hours
    * 4 - 5 hours
    * over 5 hours a day

3. f2s35a (amount of TV watched during weekdays)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 5 hours
    * over 5 hours a day
    
4. f2s35b (amount of TV watched during weekends)
    * don't watch TV
    * less than 1 hour
    * 1 - 2 hours
    * 2 - 3 hours
    * 3 - 5 hours
    * over 5 hours a day

5. bys44i (I certainly feel useless at times)
    * strongly agree
    * agree
    * disagree
    * strongly disagree
    
6. f2s66i (r feels useless at times)
    * strongly agree
    * agree
    * disagree
    * strongly disagree

Confounders (bys):
    * teachers praise my efforts (bys59h)
        * strongly agree
        * agree
        * disagree
        * strongly disagree
        
Confounders (f2)
    * f2s81c (last 30 days, # times r drank alcohol)
        * 0 occasions
        * 1-2 occcasions
        * 3-19 occasions
        * 20+ occasions
    * f2s82 (# times r had 5 drinks or more in a row)
        * none
        * once
        * twice
        * 3-5 times
        * 6-9 times
        * 10 or more times

Controls:
    * sex (sex)
    * race (race)
    * parent's highest level of education (bypared)
    * family income (byfaminc)

## Packages

```{r, message = F, warning = F}
library(here)
library(dplyr)
library(ggplot2)
library(sjlabelled)
```

## Read in data

```{r}
data <- haven::read_dta(here("class data.dta"))
```

