---

title: "Data Exploration"
output: html_document
---

## Packages

```{r, message = F, warning = F}
library(here)
library(dplyr)
library(readr)
library(broom)
library(purrr)
library(table1)
```

## Read in data

```{r}
data <-
    read_fwf(here("USSC_PV_Report_2020", "USSC_PV_Report_2020.dat"),
             fwf_widths(c(7, 1, 2,
                          1, 1, 2,
                          2, 4, 2,
                          1, 1, 1,
                          2, 3, 2,
                          2, 1, 1,
                          1, 1, 1,
                          1, 1, 4,
                          1, 2, 1,
                          10, 6, 1, 
                          6, 1, 1,
                          18, 18, 1,
                          1, 5, 4,
                          3, 3, 6,
                          1, 1),
                        c("USSCIDN", "ACCAP", "AGE", "AGGROLHI", "CAROFFAP",
                          "CIRCDIST", "COMBDRG2", "FISCALYR", "MITROLHI", "MONSEX",
                          "NEWCIT", "NEWRACE", "OFFTYPE2", "PROBATN", "PV_AGE",
                          "PV_CIRCDIST", "PV_DISP", "PV_DSHEAR", "PV_DSJANDC",
                          "PV_DSSUMMON", "PV_DSVIOL", "PV_DSWAIVER", "PV_DSWORK",
                          "PV_FY", "PV_GRADVIOL", "PV_GRADVIOL_TABLE", "PV_REVTABLE",
                          "PV_REVTABLE_CELL", "PV_SENSPLT0", "PV_SENTIMP",
                          "PV_SENTTOT", "PV_VIOLTYP", "SAFETY", "SENSPCAP",
                          "SENSPLT0", "SENTIMP", "SOURCES", "STATMAX", "STATMIN",
                          "SUPREL", "SUPRVISN", "SUPSERVED", "WEAPON", "XCRHISSR")),
             col_types = cols(PV_GRADVIOL_TABLE = "c", PV_REVTABLE_CELL = "c",
                              PV_REVTABLE = "c", PV_GRADVIOL = "c"))
```

## Some data cleaning

* Filter so we only have probation violations.
* Independent Variables:
    * Age: PV_AGE (could also use AGE which is age at original sentencing)
    * Sex: MONSEX
    * Citizenship status: NEWCIT
    * Race: NEWRACE
    * Primary offense type at the time of original sentencing: OFFTYPE2
    * Did offender admit to supervision violation?: PV_DISP
    * Grade of probation violation: PV_GRADVIOL
    * Criminal history category: XCRHISSR
* Dependent Variables:
    * Was revocation sentence above, below or within guidelines range?: PV_REVTABLE
    * Total prison sentence imposed at the revocation hearing: PV_SENTTOT (includes only prison sentences and prison sentences/alternative confinement unlike PV_SENSPLT0 which includes just alternative confinement)
    * Type of sentence imposed at revocation hearing: PV_SENTIMP

```{r}
probations <-
    data %>%
    filter(PV_VIOLTYP == 1) %>%
    select(PV_AGE, MONSEX, NEWCIT, NEWRACE, OFFTYPE2, PV_DISP, PV_GRADVIOL,
           PV_REVTABLE, PV_SENTTOT, PV_SENTIMP, XCRHISSR) %>%
    mutate(PV_AGE = as.numeric(PV_AGE),
           XCRHISSR = as.numeric(XCRHISSR),
           PV_SENTTOT = as.numeric(PV_SENTTOT),
           MONSEX = case_when(MONSEX == 0 ~ "Male",
                              MONSEX == 1 ~ "Female"),
           NEWCIT = case_when(NEWCIT == 0 ~ "U.S.A. Citizen",
                              NEWCIT == 1 ~ "Non-U.S.A. Citizen"),
           NEWRACE = case_when(NEWRACE == 1 ~ "White",
                               NEWRACE == 2 ~ "Black",
                               NEWRACE == 3 ~ "Hispanic",
                               NEWRACE == 6 ~ "Other"),
           PV_DISP = case_when(PV_DISP == 1 ~ "Admitted",
                               PV_DISP == 2 ~ "Denied",
                               PV_DISP == 3 ~ "Unclear",
                               PV_DISP == 4 ~ "Admitted and Denied"),
           PV_GRADVIOL = case_when(PV_GRADVIOL == 1 ~ "Grade A",
                                   PV_GRADVIOL == 2 ~ "Grade B",
                                   PV_GRADVIOL == 3 ~ "Grade C"),
           PV_REVTABLE = case_when(PV_REVTABLE == 0 ~ "Below Range",
                                   PV_REVTABLE == 1 ~ "Within Range",
                                   PV_REVTABLE == 2 ~ "Above Range"),
           NEWRACE = relevel(as.factor(NEWRACE), ref = "White"),
           blackIntTerm = case_when(NEWRACE == "Black" & MONSEX == "Male" ~ PV_AGE,
                                T ~ 0))
```

### Clean-up offense type category

There are a lot of offense type categories. The buckets are going to be a bit small in some cases so we need to combine them.

```{r}
offtype <-
    probations %>%
    count(OFFTYPE2) %>%
    mutate(prcnt = n / sum(n)) %>%
    arrange(desc(prcnt)) %>%
    mutate(cumsum = cumsum(prcnt))

probations <-
    probations %>%
    mutate(OFFTYPE2 = case_when(OFFTYPE2 == 27 ~ "Immigration",
                                OFFTYPE2 == 10 ~ "Drugs - Trafficking, Manufacturing, and Importing",
                                OFFTYPE2 == 18 ~ "Fraud",
                                OFFTYPE2 == 35 ~ "Traffic Violations and Other Offenses",
                                OFFTYPE2 == 17 ~ "Larceny",
                                OFFTYPE2 == 13 ~ "Firearms",
                                OFFTYPE2 == 12 ~ "Drugs - Simple Possession",
                                OFFTYPE2 == 20 ~ "Forgery/Counterfeiting",
                                OFFTYPE2 == 30 ~ "Administration of Justice",
                                OFFTYPE2 == 5 ~ "Assault",
                                T ~ "All Other Offenses"),
           OFFTYPE2 = relevel(as.factor(OFFTYPE2), ref = "Immigration"))
```

### Clean-up sentence type at revocation hearing

Combine prison only (1) and prison + alternative confinement (2) into one category (1). All other revocation decisions which do not include a prison sentence are combined (0).

```{r}
probations <-
    probations %>%
    mutate(PV_SENTIMP =
               case_when(PV_SENTIMP == 1 ~ 1,
                         PV_SENTIMP == 2 ~ 1,
                         T ~ 0))
```

### Create descriptive table for probations

```{r}
table1::label(probations$PV_AGE) <- "Age"
table1::label(probations$MONSEX) <- "Sex"
table1::label(probations$NEWCIT) <- "Citizenship Status"
table1::label(probations$NEWRACE) <- "Race"
table1::label(probations$OFFTYPE2) <- "Offense Type"
table1::label(probations$PV_DISP) <- "Probationer Admittance"
table1::label(probations$PV_GRADVIOL) <- "Probation Violation"
table1::label(probations$PV_SENTIMP) <- "Revocation Decision"
table1::label(probations$XCRHISSR) <- "Federal Criminal History Score"

descrptTable <-
    table1(~PV_AGE + MONSEX + NEWCIT + NEWRACE + OFFTYPE2 + PV_DISP + PV_GRADVIOL +
               as.factor(PV_SENTIMP) + as.factor(XCRHISSR),
       data = probations)
```

## Run Analyses

Logistic regression as to whether or not an offender received a prison sentence or not.

```{r}
prisonOrNot <-
    probations %>%
    select(-PV_SENTTOT, -PV_REVTABLE) %>%
    filter(across(everything(), ~!is.na(.x)))

prisonOrNotC <-
    probations %>%
    filter(PV_GRADVIOL == "Grade C") %>%
    select(-PV_SENTTOT, -PV_REVTABLE, -PV_GRADVIOL) %>%
    filter(across(everything(), ~!is.na(.x)))

prisonOrNotAll <-
    probations %>%
    select(-PV_SENTTOT, -PV_REVTABLE, -PV_GRADVIOL, -XCRHISSR) %>%
    filter(across(everything(), ~!is.na(.x)))

prisonOrNotMissings <-
    probations %>%
    mutate(PV_GRADVIOL = if_else(is.na(PV_GRADVIOL), "Missing", PV_GRADVIOL)) %>%
    select(-PV_SENTTOT, -PV_REVTABLE) %>%
    filter(across(everything(), ~!is.na(.x)))

logitModel <- glm(PV_SENTIMP ~., data = prisonOrNot, family = "binomial")
logitMissing <- glm(PV_SENTIMP ~., data = prisonOrNotMissings, family = "binomial")
summary(logitModel)
summary(logitMissing)

logitModelC <- glm(PV_SENTIMP ~., data = prisonOrNotC, family = "binomial")
summary(logitModelC)

logitModelAll <- glm(PV_SENTIMP ~., data = prisonOrNotAll, family = "binomial")
summary(logitModelAll)
```

OLS regression for sentence length conditional on receiving a prison sentence.

```{r}
revocationSentenceLength <-
    probations %>%
    select(-PV_REVTABLE, -PV_SENTIMP) %>%
    filter(across(everything(), ~!is.na(.x)))

olsModel <- lm(log(PV_SENTTOT) ~ ., data = revocationSentenceLength)
summary(olsModel)
```

Multinomial Regression on Guideline Departures

```{r}
departures <-
    probations %>%
    select(-PV_SENTIMP, -PV_SENTTOT) %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(PV_REVTABLE = as.factor(PV_REVTABLE))

multinomialModels <-
    map(list("Above Range", "Below Range", "Within Range"),
        function(df, level) {
            df$PV_REVTABLE <- relevel(df$PV_REVTABLE, ref = level)
            nnet::multinom(PV_REVTABLE ~ ., data = df)},
        df = departures)

map(multinomialModels, summary)

multinomialDf <- tidy(multinomialModels[[2]], exponentiate = T)
```

Logsitic Regressions for Guideline Departures

```{r}
withinVsDepart <-
    probations %>%
    select(-PV_SENTIMP, -PV_SENTTOT) %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(PV_REVTABLE = case_when(PV_REVTABLE == "Within Range" ~ 1,
                                   PV_REVTABLE == "Below Range" ~ 0,
                                   PV_REVTABLE == "Above Range" ~ 0))

logitModel2 <- glm(PV_REVTABLE ~., data = withinVsDepart, family = "binomial")
summary(logitModel2)
```

```{r}
BelowVsAbove <-
    probations %>%
    select(-PV_SENTIMP, -PV_SENTTOT) %>%
    filter(across(everything(), ~!is.na(.x))) %>%
    mutate(PV_REVTABLE = case_when(PV_REVTABLE == "Within Range" ~ NA_real_,
                                   PV_REVTABLE == "Below Range" ~ 0,
                                   PV_REVTABLE == "Above Range" ~ 1))

logitModel3 <- glm(PV_REVTABLE ~., data = BelowVsAbove, family = "binomial")
summary(logitModel3)
```

